Security Vulnerabilities of Container Images in Neuroimaging

### 2020-09-09 paper.tex
Major edits in paper, reimplemented analysis in Jupyter

\journal{gigascience}
%%%% Packages %%%%
\usepackage{siunitx}
\usepackage{algpseudocode} % Algorithmic environment
\usepackage{xspace}
\usepackage{csvsimple}
\usepackage{minted} % Used for JSON highlighting
\usepackage{datatool}
\usepackage{booktabs}
\usepackage{xcolor}
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage{multirow}
\usepackage{array}
\usepackage{algorithm}
\usepackage{caption}
\usepackage[justification=centering]{caption}
\usepackage{subcaption}
\usepackage{listings}
\usepackage{verbatim}
\usepackage{adjustbox}
\usepackage{makecell}
\usepackage{titlesec}
\usepackage{anyfontsize}
\usepackage[flushleft]{threeparttable}

%%%% Commands %%%%
\newcommand{\todo}[1]{\color{red}\textbf{TODO:}#1\color{black}}
\newcommand{\change}[2]{\color{cyan}Changes: #1\color{black}}
\newcommand{\reprozip}[0]{ReproZip}
\newcommand{\rom}[1]{\lowercase\expandafter{\romannumeral #1\relax}}
\newcommand{\tristan}[1]{\color{blue}From Tristan: #1\color{black}}

\renewcommand{\labelitemi}{$\textendash$}
\title{Vulnerability Analysis of Container Images in Neuroimaging}
\begin{filecontents*}{resul.csv}
labels,names,A,C,V,Stools
a,bids/example:latest,838,663,683,
b,bids/freesurfer:latest,353,215,192,
c,bids/ndmg:latest,934,757,782,
d,bids/broccoli:v1.0.0,203,203,117,
e,bids/fibredensityandcrosssection:latest,1086,849,868,
f,bids/spm:latest,191,246,235,
g,poldracklab/mriqc:0.15.0,432,328,332,
h,bids/cpac:v1.0.1a\_22,698,524,535,
I,bids/hyperalignment:latest,1717,1327,1433,
j,bids/mindboggle:0.0.4-1,671,955,569,
k,bids/mrtrix3\_connectome:latest,275,157,158,
l,bids/rs\_signal\_extract:0.1,867,454,473,
m,bids/aa:v0.2.0,189,244,233,
n,bids/niak:latest,1482,1366,1420,
o,bids/oppni:v0.7.0-1,913,698,723,
p,poldracklab/fmriprep:unstable,516,355,393,
q,bids/brainiak-srm:latest,897,484,503,
r,bids/nipypelines:0.3.0,641,878,451,
s,bids/hcppipelines:v3.17.0-18,526,873,620,
t,bids/magetbrain:latest,112,117,126,
u,bids/tracula:v6.0.0-4,874,664,683,
v,bids/baracus:latest,172,238,216,
w,bids/antscorticalthickness:v2.2.0-1,289,27,no data,
x,bids/dparsf:v4.3.12,507,750,757,
y,bids/afni\_proc:latest,67,16,no data,
z,bids/rshrf:1.0.1,5,3,3,
aa,bids/ndmg:v0.1.0,934,758,782,
ab,mcin/qeeg:latest,121,121,142,
ac,camarasu/creaphase:0.3,148,148,186,
ad,bigdatalabteam/hcp-prefreesurfer:exec-centos7.freesurferbuild-centos4-latest:latest,257,266,157,
ae,mcin/docker-fsl:latest,178,178,201,
af,gkiar/onevoxel:v0.3.0rc2,5,2,2,
ag,bigdatalabteam/hcp-prefreesurfer:exec-centos7-fslbuild-centos5-latest:latest,257,266,157,
ah,bt5e/ants:latest,592,600,336,
ai,poldracklab/fmriprep:1.2.3,591,441,447,
aj,gkiar/dwipreproc\_fsl-5.0.11\_minified:latest,391,258,263,
ak,gkiar/mask2boundary:v0.1.0,15,11,11,
al,mcin/ica-aroma:latest,178,178,202,
am,c3genomics/genpipes,,,,99
an,bioinformatics-group/aqua-singularity-recipe,,,,7
ao,MontrealSergiy/BEst,,,,106
ap,shub://aces/cbrain-containers-recipes:fsl\_v6.0.1,,,,206
aq,shots47s/bids-freesurfer-6.0,,,,212
ar,shots47s/bids-fmriprep-1.2.3,,,,453
\end{filecontents*}

@@

% Containers are popular on HPC, in particular Singularity
Software containers have emerged has an efficient solution to deploy
scientific data analyses on High Performance Computing (HPC) clusters, due
to their portability, ease of use, and limited overhead. The
Singularity~\cite{kurtzer2017singularity} framework is often preferred to
Docker\footnote{\url{http://docker.com}} on HPC clusters, due to its
secure handling of multi-user environments, and its convenient support for
Docker images. Singularity is now available in dozens of HPC
clusters around the globe, and used routinely for Big Data analyses.

% Containers are meant to isolate applications from the host, but security remains an issue
Using core features of the Linux kernel such as namespaces, control groups
and chroot, containers isolate processes from the
host computer,  
and control the memory, CPU, network and file-system resources assigned to
them. However, containers still share the kernel, mounted file systems and
some devices with the host, which raises potential security issues. In
particular, privilege escalation, denial of service and information leak
are always plausible~\cite{gantikow2016providing}. Therefore, security remains 
an important concern when deploying containers~\cite{bettini2015vulnerability}.

% Container images, our study
This study is a vulnerability analysis of container images used in HPC
clusters. Container images contain complete operating-system distributions, including
the data analysis software and its required dependencies.
Images are rarely updated, as they are often built once
for the entire duration of an analysis, to avoid results perturbations
coming from software updates~\cite{gronenschild2012effects, glatard2015reproducibility}. Images also
commonly include more dependencies than required by a specific analysis, to 
enable reusability across experiments.

Our research questions are the following:

@@

\textit{RQ3: Can the amount of vulnerabilities be reduced by minimizing the images?} 
Container images often include more software packages than required for a
typical analysis. We report on the impact of this additional software on
the presence of vulnerabilities.

We answer these questions taking the particular example of neuroimaging, a
domain that has entered the Big Data era as many other
ones~\cite{van2014human}. We focus on the analysis of container images
present in BIDS apps~\cite{gorgolewski2017bids} and Boutiques~\cite{glatard2018boutiques}, two containerized
application frameworks commonly used in neuroimaging.

The next section describes the materials and tools used in the experiments,
including container images and image scanners. We then describe our image
update and minimization methodology to reduce vulnerabilities. Results review
the detected vulnerabilities in container images, explain observed
discrepancies between scanners, and report on the efficiency of image
update and size reduction. As a conclusion, we provide a set of image
creation guidelines for a secure deployment on HPC clusters.

The existing research work on containers focuses mainly on the security of containers.
This focus is justified by the fact that containers expose the host's resources
(e.g., file system/ IPC) to the guest system. This feature raises a confidentiality
threat for the applications running on the same host. Previous studies evaluated
the security of Docker engine ~\cite{martin2018docker, sultan2019container, combe2016docker, bui2015analysis}.
Other work studied security vulnerabilities in Docker images. A study in 2015 by Gummaraju et al.
states that over 30\% of official images in docker hub contain high priority 
security vulnerabilities~\cite{gummaraju2015over}. Shu et al. show that both official and community images 
contain more than 180 vulnerabilities on an average~\cite{Shu2017}.
Zerouali et al. studied that 
outdated packages lead to bugs and severity vulnerabilities in Docker
images~\cite{zerouali2019relation}.

@@

Since the aim of this research work is to detect security vulnerabilities
in container images and attempt to examine the effect of different
manipulations on these images, it is needed to have a representative set of
images. Additionally, for the purpose of detecting vulnerabilities,
utilization of image scanners is required. We utilize container
images from two popular application frameworks, as well as
four of the main image scanners.

@@

We scan all container images, available at the time of this study, on two containerization frameworks
used in neuroscience: BIDS
apps~\cite{gorgolewski2017bids} and Boutiques~\cite{glatard2018boutiques}, totalling
44 container images.
Out of these 44 images, 26 were obtained from BIDS apps
(\url{https://bids-apps.neuroimaging.io/apps/}) 
and 18 from Boutiques. In fact, at the time of the study, BIDS apps had 27 images,
out of which one wasn't available on DockerHub. Boutiques had 49 images,
%listed using Boutiques \texttt{bosh search} command,
however, only 23 unique images were listed, out of which 3 couldn't be retrieved and 2
were already included in BIDS apps. All the final 26 images
from BIDS apps are Docker images, whereas Boutiques contains 18 Docker images
and 6 Singularity images.

@@

We use four popular image scanners, namely Anchore~\cite{github_2019},
Vuls~\cite{future-architect_2019}, Clair~\cite{arminc_2019} and
Singularity Container Tools --- Stools~\cite{stools}. Anchore, Vuls and Clair support Docker
images while Stools support Singularity. Different vulnerabilities may be
reported by scanners if scanning experiments take place on different dates.
To avoid such discrepancies, we freeze the vulnerability
databases used by these scanners as of 2019-09-25, when we initiate our
experiments. This allows for a fair comparison between the scanners.

\subsubsection{Anchore}

Anchore is an end-to-end, open-source container security platform. It
analyze	container images and lists OS
packages, non-OS packages (Python, Java, Gem, and Npm), and files.
In our experiments, we use Anchore Engine version 0.5.0 through Docker image \texttt{anchore/anchore-engine:v0.5.0}, and
Anchore vulnerability database version 0.0.11.

Anchore compares package versions, which are present inside the
image, with the vulnerability database in order to flag vulnerabilities.
To get a vulnerability feed of various OS distributions, Anchore refers to
different sources. Figure~\ref{database} displays Anchore's vulnerability sources for
Ubuntu, Debian, Alpine, and Centos, which are present inside taken images.
Anchore also refers to National
Vulnerability Database (NVD) to get metadata of vulnerabilities.

@@

\subsubsection{Vuls}

Vuls is an open-source vulnerability scanner for Linux and FreeBSD. It
offers both static and dynamic scanning, and both local and remote
scanning. In our experiments, we use remote dynamic scanning, that is, Vuls
connects to the running containers that needs to be scanned through SSH, and queries the package manager
in the container. We use Vuls 0.9.0, executed through Docker image
\texttt{vuls/vuls:0.9.0}. At the time of this study, Vuls gets vulnerability feed
from a configured set of sources. Figure~\ref{database} lists vulnerability databases
referred by Vuls for scanned four OS distributions i.e. Alpine, Ubuntu, Debian, and Centos.
Apart from this, like Anchore, Vuls also refers NVD database for getting vulnerability metadata.
Figure~\ref{database} shows that Vuls
refers to Open Vulnerability and Assessment Language (OVAL) database for Ubuntu, Debian, and Centos.
It is worth noting that there is
no vulnerability data for Ubuntu 17.04 and 17.10 distributions in OVAL database, meaning that images
with these distributions cannot be scanned with Vuls.

\subsubsection{Clair}

Clair (\url{https://github.com/quay/clair}) is an open-source and extensible
vulnerability scanner for Docker and appc container images, developed by
CoreOS (now Container Linux), a Linux distribution to deploy container
clusters. Clair has a client-server architecture, in which the server
scans Docker images layer by layer and maintains a database of vulnerabilities. We use Clair through Clair-scanner
(\url{https://github.com/arminc/clair-scanner}), a tool to facilitate the testing
of container images against a local Clair server. Clair-scanner scans the image,
prepares a list of vulnerabilities, compares that list against a
whitelist, and flags vulnerabilities that are not present in the whitelist.
In our experiments, we did not use a whitelist to filter scanning results in
order to make a fair comparison between scanners.
Clair-scanner maintains a Docker image with the up-to-date vulnerability
database from a set of different sources. Figure~\ref{database} lists
vulnerability databases that Clair referred for Alpine, Ubuntu, Debian, and Centos.
Clair also uses NVD database for getting vulnerability metadata.
We use Clair version 2.0.6, executed through
Docker image \texttt{arminc/clair-local-scan:v2.0.6}. For the vulnerability
database, we use Docker image \texttt{arminc/clair-db:latest}, last
updated on 2019-09-18.

\subsubsection{Singularity Tools}

Singularity Tools (Stools)
are an extension of Clair for Singularity images. Stools
exports Singularity images to \texttt{tar.gz} format, acting as a single layer Docker image
to circumvent the Docker-specific requirements in the Clair API.
In our experiments, we use Singularity Tools version 3.2.1 through Docker
image
\texttt{vanessa/stools-clair:v3.2.1}, which is using another Docker
image \texttt{arminc/clair-db:latest} for referring to vulnerability feed.
Since Stools uses Clair internally for scanning, vulnerability databases used
by Stools are same as mentioned for Clair.
To scan Singularity images, we follow the steps mentioned in
\href{https://github.com/singularityhub/stools}{Stools Github}.

\subsection{Ubuntu Vulnerability Database}

\change{There are two types of vulnerability databases that are referred
by the scanners in these experiments. First one is Open Vulnerability and Assessment Language (OVAL)
database, which is an international open standard and a common means of determining the existence of
vulnerabilities in softwares. OVAL supports various OS distributions e.g. Ubuntu, Debian, Centos, etc.
Second type of databases used by scanners are vulnerability databases from specific OS vendors e.g. Alpine-SecDB
from Alpine,
Debian Bug Tracker from Debian, Ubuntu CVE Tracker from Ubuntu, etc., where they
keep a track of their OS-specific vulnerabilities by assigning a status.
%In particular, there is a discrepancy in the way Anchore and Clair
%interpret Ubuntu CVE Tracker.
Out of four distributions that are present in our images Vuls uses
OVAL database for Debian, Ubuntu, Centos and refers to Alpine-SecDB
database for Alpine as this distribution is currently not supported by OVAL.
Clair and Anchore mostly refers to vulnerability databases provided by specific OS
vendors except in one case, where Anchore refers to OVAL i.e. Centos.

The major discrepancies are shown by Anchore and Clair in Ubuntu. Both refers to Ubuntu CVE Tracker database for Ubuntu images.
The way in which both scanners interpret this database is different.
%
%There are different vulnerability databases, which have
%different formats of data.
%Out of which Open Vulnerability and Assessment Language (OVAL) is an international
%open standard and a common
%means of promoting publicly available security issues. 
%It provides details about software vulnerabilities. It helps to determine
%whether vulnerabilities exist in local machines or not. Apart from OVAL, operating system
%vendors also keep track of their vulnerabilities separately and assign
%a status to each vulnerability.
%For example, Ubuntu maintains
%Ubuntu CVE Tracker, Debian maintains Debian Security Bug Tracker, etc.
%Each vulnerability is assigned a unique Common Vulnerabilities and Exposures (CVE) number
%by the MITRE corporation, who maintains a list of all vulnerabilities of all distributions.
%In this paper, we use vulnerability and CVE interchangeably.
%Ubuntu CVE Tracker is used by Anchore and Clair in our experiments
%and there is discrepancies in the way both scanners interpret results.
%
%BIDS apps and Boutiques images in our experiments have four Linux
%distributions i.e. Ubuntu, CentOS, Debian, and Alpine. 
%Out of these distributions, Ubuntu assigns a vulnerability status to
%all vulnerabilities before keeping them in its vulnerability database.
%Each vulnerability is assigned a unique Common Vulnerabilities and Exposures (CVE) number
%by the MITRE corporation, who maintains a list of all vulnerabilities of all distributions.
%Ubuntu keeps a record of Ubuntu-specific vulnerabilities in Ubuntu CVE Tracker using their CVE
%numbers. In this paper, we use vulnerability and CVE interchangeably.
%The vulnerability status in Ubuntu CVE Tracker is subjected to change
%depending on the work progress to fix that vulnerability.
%
%Regarding Ubuntu, it has different
%types of releases e.g. Regular, Long Term Support (LTS), and Extended Security Maintenance (ESM) release.
%Ubuntu maintains its vulnerability data in Ubuntu CVE tracker, which is referred by Anchore and Clair.
%There is another vulnerability database i.e. Open Vulnerability and Assessment Language (OVAL) repository,
%which is maintained by MITRE corporation. Vuls refers to OVAL repository for getting Ubuntu vulnerability feed.
%The vulnerability data format of both databases is different. OVAL keeps a separate xml file for each release,
%which contains that release specific vulnerability information.
Ubuntu CVE Tracker keeps a separate file for each vulnerability, which
has an entry for all Ubuntu releases combined with the current status
of that vulnerability in that release.
In particular, Clair and Anchore show discrepancies for Ubuntu 14.04 which have
two entries for each vulnerability, one for Ubuntu 14.04 Long Term Support (LTS)
and second for Ubuntu 14.04 Extended Security Maintenance (ESM) release. As Ubuntu 14.04 LTS
reached end-of-life on 2019-04-30, so Ubuntu ignores new vulnerabilities in LTS release and
fix them only in ESM release. Consequently, a same vulnerability may have different vulnerability status in LTS and ESM releases,
which is causing discrepancies. This is because Anchore refers to Ubuntu 14.04 LTS, whereas Clair refers to
ESM.
%Further, in Ubuntu a particular release can be of different types i.e. Regular, Long Term Support (LTS), and Extended Security Maintenance (ESM) release. 
%Through our experiments, we find that Anchore and Clair report Ubuntu vulnerabilities 
%differently depending on either the different Ubuntu release type or vulnerability status. 
It is important to provide an overview of these Ubuntu vulnerability statuses, 
since Ubuntu is heavily utilized in our experiments. 
%Ubuntu's Regular release is supported only for 9 months, whereas LTS is supported for 5 years. Once LTS support
%reaches End-Of-Life (EOL), Ubuntu decides whether to provide extended security support or not, which comes
%in the form of ESM release.
%ESM is a paid service that can be requested from Ubuntu to get security 
%updates even after the EOL of a particular Ubuntu release. ESM and LTS releases of a particular Ubuntu version have different vulnerability
%databases, because after EOL of a release, a vulnerability is fixed only in packages of ESM; however, it still
%exists in LTS.
%
%As an example, Ubuntu has ended support for some releases, such as 14.04 LTS,
%17.04, and 17.10, as they reached EOL. However, Ubuntu decided to provide optional extended support for Ubuntu 14.04.
%To use this extended support, users have to purchase Ubuntu 14.04 ESM release.
%The images in our experiments have Ubuntu 14.04 LTS, so scanners should refer
%to LTS release only while collecting data from the vulnerability databases.
%More interestingly, Clair refers to ESM release vulnerability database, however, this may lead to discrepancies in results.
%Further, if an image is using a Linux release where EOL has been reached that image may
%turns to be vulnerable as it is not possible to retrieve any updates.
%
In Ubuntu CVE Tracker, for a vulnerability all Ubuntu releases are listed with the package name
in which that vulnerability exists and are entitled a vulnerability status
(\url{https://git.launchpad.net/ubuntu-cve-tracker/plain/README}), which is
encoded in the following form:
\newline \\
\noindent <release>\_<source-package>: <status> (<version/notes>)
\newline\\
The <version/notes> given with a vulnerability status provide more

>>>>

\journal{gigascience}
%%%% Packages %%%%
\usepackage{siunitx}
\usepackage{algpseudocode} % Algorithmic environment
\usepackage{xspace}
\usepackage{csvsimple}
\usepackage{minted} % Used for JSON highlighting
\usepackage{datatool}
\usepackage{booktabs}
\usepackage{xcolor}
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage{multirow}
\usepackage{array}
\usepackage{algorithm}
\usepackage{caption}
\usepackage[justification=centering]{caption}
\usepackage{subcaption}
\usepackage{listings}
\usepackage{verbatim}
\usepackage{adjustbox}
\usepackage{makecell}
\usepackage{titlesec}
\usepackage{anyfontsize}
\usepackage[flushleft]{threeparttable}
\usepackage{xspace}

%%%% Commands %%%%
\newcommand{\todo}[1]{\color{red}\textbf{TODO:}#1\color{black}}
\newcommand{\change}[2]{\color{cyan}Changes: #1\color{black}}
\newcommand{\reprozip}[0]{ReproZip\xspace}
\newcommand{\rom}[1]{\lowercase\expandafter{\romannumeral #1\relax}}
\newcommand{\tristan}[1]{\color{blue}From Tristan: #1\color{black}}

\renewcommand{\labelitemi}{$\textendash$}
\title{Security Vulnerabilities of Container Images in Neuroimaging}
\begin{filecontents*}{resul.csv}
labels,names,A,C,V,Stools
a,bids/example:latest,838,663,683,
b,bids/freesurfer:latest,353,215,192,
c,bids/ndmg:latest,934,757,782,
d,bids/broccoli:v1.0.0,203,203,117,
e,bids/fibredensityandcrosssection:latest,1086,849,868,
f,bids/spm:latest,191,246,235,
g,poldracklab/mriqc:0.15.0,432,328,332,
h,bids/cpac:v1.0.1a\_22,698,524,535,
I,bids/hyperalignment:latest,1717,1327,1433,
j,bids/mindboggle:0.0.4-1,671,955,569,
k,bids/mrtrix3\_connectome:latest,275,157,158,
l,bids/rs\_signal\_extract:0.1,867,454,473,
m,bids/aa:v0.2.0,189,244,233,
n,bids/niak:latest,1482,1366,1420,
o,bids/oppni:v0.7.0-1,913,698,723,
p,poldracklab/fmriprep:unstable,516,355,393,
q,bids/brainiak-srm:latest,897,484,503,
r,bids/nipypelines:0.3.0,641,878,451,
s,bids/hcppipelines:v3.17.0-18,526,873,620,
t,bids/magetbrain:latest,112,117,126,
u,bids/tracula:v6.0.0-4,874,664,683,
v,bids/baracus:latest,172,238,216,
w,bids/antscorticalthickness:v2.2.0-1,289,27,no data,
x,bids/dparsf:v4.3.12,507,750,757,
y,bids/afni\_proc:latest,67,16,no data,
z,bids/rshrf:1.0.1,5,3,3,
aa,bids/ndmg:v0.1.0,934,758,782,
ab,mcin/qeeg:latest,121,121,142,
ac,camarasu/creaphase:0.3,148,148,186,
ad,bigdatalabteam/hcp-prefreesurfer:exec-centos7.freesurferbuild-centos4-latest:latest,257,266,157,
ae,mcin/docker-fsl:latest,178,178,201,
af,gkiar/onevoxel:v0.3.0rc2,5,2,2,
ag,bigdatalabteam/hcp-prefreesurfer:exec-centos7-fslbuild-centos5-latest:latest,257,266,157,
ah,bt5e/ants:latest,592,600,336,
ai,poldracklab/fmriprep:1.2.3,591,441,447,
aj,gkiar/dwipreproc\_fsl-5.0.11\_minified:latest,391,258,263,
ak,gkiar/mask2boundary:v0.1.0,15,11,11,
al,mcin/ica-aroma:latest,178,178,202,
am,c3genomics/genpipes,,,,99
an,bioinformatics-group/aqua-singularity-recipe,,,,7
ao,MontrealSergiy/BEst,,,,106
ap,shub://aces/cbrain-containers-recipes:fsl\_v6.0.1,,,,206
aq,shots47s/bids-freesurfer-6.0,,,,212
ar,shots47s/bids-fmriprep-1.2.3,,,,453
\end{filecontents*}

@@

% Containers are popular on HPC, in particular Singularity
Software containers have emerged has an efficient solution to deploy
scientific data analyses on High Performance Computing (HPC) clusters, due
to their portability, ease of use, and limited overhead. The
Singularity~\cite{kurtzer2017singularity} framework is often preferred to
Docker\footnote{\url{http://docker.com}} for HPC deployments, due to its
secure handling of multi-user environments, and its convenient support for
Docker images. Singularity is now available in dozens of HPC
clusters around the globe, and used routinely for Big Data analyses.

% Containers are meant to isolate applications from the host, but security
% remains an issue
Leveraging core features of the Linux kernel such as namespaces, control
groups and chroot, containers isolate processes from the host computer,  
and control the memory, CPU, network and file-system resources assigned to
them. Nevertheless, containers still share the kernel, mounted file systems
and some devices with the host, which raises security
issues~\cite{martin2018docker, sultan2019container, combe2016docker} and
opens the door to privilege escalation, denial of service, information
leak and other types of attacks~\cite{gantikow2016providing}.

% Container images, our study
Container images often include nearly complete operating-system
distributions in addition to the data analysis software and its required
dependencies. Images are rarely updated, as they are often built once for
the entire duration of an analysis, to avoid results perturbations due to
software updates~\cite{gronenschild2012effects,
glatard2015reproducibility}. Images also commonly include more dependencies
than required by a specific analysis, to facilitate reusability across
experiments. As a result, it has been shown that over 30\% of official
images in DockerHub contain high priority security
vulnerabilities~\cite{gummaraju2015over}, that both official and community
images contain more than 180 vulnerabilities on average~\cite{Shu2017}, and
that vulnerabilities are often due to outdated
packages~\cite{zerouali2019relation}.

In this study, we focus on the vulnerability analysis of container images
used in HPC clusters.  
Our research questions are the following:

@@

\textit{RQ3: Can the amount of vulnerabilities be reduced by minifying the images?} 
Container images often include more software packages than required for a
typical analysis. We report on the impact of this additional software on
the presence of vulnerabilities.

We answer these questions by taking the particular example of neuroimaging,
a domain that has entered the Big Data era as many other
ones~\cite{van2014human}. The next section describes the materials and
tools used in the experiments, including container images and image
scanners. We then describe our image update and minification methodology to
reduce vulnerabilities. Results present the detected vulnerabilities in
container images, explain observed discrepancies between scanners, and
report on the efficiency of image update and size reduction. As a
conclusion, we provide a set of image creation guidelines for a more secure
deployment on HPC clusters.

% We focus on the analysis of container images
% present in BIDS apps~\cite{gorgolewski2017bids} and Boutiques~\cite{glatard2018boutiques}, two containerized
% application frameworks commonly used in neuroimaging.

@@

To detect security vulnerabilities
in container images and examine the effect of different
manipulations on these images, we used container
images from two popular application frameworks, as well as
four of the main image scanners.

@@

We scanned all container images available at the time of this study on two containerization frameworks
used in neuroscience: BIDS
apps~\cite{gorgolewski2017bids} and Boutiques~\cite{glatard2018boutiques},
totalling
44 container images.
Out of these 44 images, 26 were obtained from BIDS apps 
and 18 from Boutiques. At the time of the study, BIDS apps had 27 images,
out of which one wasn't available on DockerHub. Boutiques had 49 images,
%listed using Boutiques \texttt{bosh search} command,
however, only 23 unique images were listed, out of which 3 couldn't be retrieved and 2
were already included in BIDS apps. All the final 26 images
from BIDS apps were Docker images, whereas Boutiques contained 18 Docker images
and 6 Singularity images.

@@

We used four popular image scanners, namely Anchore~\cite{github_2019},
Vuls~\cite{future-architect_2019}, Clair~\cite{arminc_2019} and
Singularity Container Tools --- Stools~\cite{stools}. Anchore, Vuls and Clair support Docker
images while Stools support Singularity.

Anchore is an end-to-end, open-source container security platform. It
analyzes container images and lists OS
packages, non-OS packages (Python, Java, Gem, and Npm), and files.
In our experiments, we use Anchore Engine version 0.5.0 through Docker image \texttt{anchore/anchore-engine:v0.5.0}, and
Anchore vulnerability database version 0.0.11.

@@

Vuls is an open-source vulnerability scanner for Linux and FreeBSD. It
offers both static and dynamic scanning, and both local and remote
scanning. In our experiments,we used Vuls 0.9.0, executed through Docker image
\texttt{vuls/vuls:0.9.0} in remote dynamic mode.

Clair (\url{https://github.com/quay/clair}) is an open-source and extensible
vulnerability scanner for Docker and appc container images, developed by
CoreOS (now Container Linux), a Linux distribution to deploy container
clusters. Clair has a client-server architecture, in which the server
scans Docker images layer by layer and maintains a database of vulnerabilities. We used Clair through Clair-scanner
(\url{https://github.com/arminc/clair-scanner}), a tool to facilitate the testing
of container images against a local Clair server. 
% Clair-scanner scans the image,
% prepares a list of vulnerabilities, compares that list against a
% whitelist, and flags vulnerabilities that are not present in the whitelist.
% In our experiments, we did not use a whitelist to filter scanning results in
% order to make a fair comparison between scanners.
% Clair-scanner maintains a Docker image with the up-to-date vulnerability
% database from a set of different sources.
% Figure~\ref{database} lists
% vulnerability databases that Clair referred for Alpine, Ubuntu, Debian, and Centos.
% Clair also uses NVD database for getting vulnerability metadata.
We used Clair version 2.0.6, executed through
Docker image \texttt{arminc/clair-local-scan:v2.0.6}. For the vulnerability
database, we use Docker image \texttt{arminc/clair-db:latest}, last
updated on 2019-09-18.

Singularity Tools (Stools)
are an extension of Clair for Singularity images. Stools
exports Singularity images to \texttt{tar.gz} format, acting as a single layer Docker image
to circumvent the Docker-specific requirements in the Clair API.
In our experiments, we used Singularity Tools version 3.2.1 through Docker
image
\texttt{vanessa/stools-clair:v3.2.1}.
% , which is using another Docker
% image \texttt{arminc/clair-db:latest} for referring to vulnerability feed.
Since Stools uses Clair internally for scanning, vulnerability databases used
by Stools are same as mentioned for Clair.
To scan Singularity images, we followed the steps mentioned in
\href{https://github.com/singularityhub/stools}{Stools Github}.

\subsection{Vulnerability Databases}

Scanners refer to two types of vulnerability databases. The first one is
the Open Vulnerability and Assessment Language (OVAL) database, an
international open standard that supports various OS distributions
including Ubuntu, Debian and Centos but not Alpine. The second one are vulnerability
databases from specific OS distributions, such as Alpine-SecDB, Debian
Security Bug Tracker, Ubuntu CVE Tracker, or Red Hat Security Data. OS
distributions often assign a status to each vulnerability, to keep track of
required and available security fixes in different versions of the
distribution. Figure~\ref{database} shows the vulnerabilities databases used by each scanner 
by OS distribution. Vuls uses OVAL databases for all distributions except Alpine, since Alpine is not available in OVAL. On the contrary, 
Clair exclusively refers to distribution-specific databases. Anchore uses OVAL only for CentOS \tristan{why?}. It is also worth noting that there is
no vulnerability data for Ubuntu 17.04 and 17.10 distributions in the OVAL database, since these distributions have reached end of life, meaning that images
with these distributions cannot be scanned with Vuls \tristan{why doesn't vuls use Ubuntu CVE tracker?}.
\tristan{In Vuls, what happens when 2 databases are used?}

Anchore also refers to National Vulnerability Database (NVD) to get
metadata of vulnerabilities. Apart from this, like Anchore, Vuls also
refers NVD database for getting vulnerability metadata \tristan{same
question}. \tristan{why is it not in the table?}.

For Centos images, Anchore and Clair gives scanning results using Red Hat
Security Advisory (RHSA) identifiers for Centos images, whereas Vuls uses
CVE identifiers. RHSA identifiers specifies security patches that are
pending to apply and CVE identifiers specifies vulnerabilities that exist
in the system. We mapped RHSA identifiers to corresponding CVE identifiers,
to allow for a comparison between scanners.

Different vulnerabilities may be
reported by scanners if scanning experiments take place on different dates.
To avoid such discrepancies, we froze the vulnerability
databases used by these scanners as of 2019-09-25, when we initiated our
experiments.\tristan{this contradicts the date used for Clair}

\subsection{Image Update}

A first approach to reduce the number of vulnerabilities in container
images is to update their packages to the latest version packaged by the OS
distribution. To study the effect of such updates, we developed a script
(available
\href{https://github.com/big-data-lab-team/container-vulnerabilities-paper/blob/master/Scripts/update}{here})
to identify the package manager in the image, and invoke it to update all
OS packages. \tristan{what was the date at which update was run?}

\subsection{Image Minification}

%We revert to image minimization to reduce the number of vulnerabilities without affecting results reproducibility.
%An interesting choice to do so would be to use 
%Neurodocker~\cite{neurodocker}.
%This tool traces and deletes unnecessary files.
%However, it does not affect the number of installed packages.
%Since image scanners report vulnerabilities according to the installed
%packages. Therefore, after
%minimization the same number of vulnerabilities is still reported.
%Consequently, we need to reduce the number of installed packages
%in the images by deleting unnecessary packages. For that, it is required to figure
%out which packages we need for the pipeline to function properly.

A second approach to reduce the number of vulnerabilities in the images is
to remove unnecessary packages, an operation potentially specific to each
analysis. We used the open-source \reprozip tool~\cite{rampin2016reprozip}
to capture the list of packages used by an analysis. For this purpose,
\reprozip first captures the list of files involved in the analysis,
through system call interception, then retrieves the list of associated
software packages, by querying the package manager. We extend this list
with a passlist of packages required for the system to function
\tristan{can you give an example?}, and with all the dependencies of the
required packages, retrieved using Debtree~\cite{debtree} (Repoquery
(\url{https://linux.die.net/man/1/repoquery}) could be used in RPM-based
distributions instead). Our minification script, available
\href{https://github.com/big-data-lab-team/container-vulnerabilities-paper/tree/master/Scripts/minification}{here},
installs \reprozip in the image to minify, runs an analysis to collect a
\reprozip trace, and finally deletes all unnecessary packages.

Using this approach, we minified five Debian- or Ubuntu-based BIDS app images,
using basic analysis examples found in the applications documentation.

% \subsection{Hybrid Approach}

% Following the two previous approaches, we employ a hybrid approach, where
% we combine both update and size reduction processes in an attempt to
% lessen number of vulnerabilities or to test the effect of
% combining two approaches. To allow for that, we first update all packages in
% the image, then delete all unnecessary packages. Applying these processes in reverse
% may reduce the needed work; however this is not possible in all cases. For instance,
% some packages, such as \texttt{apt-transport-https}, are required to fetch updates, deleting these packages
% ahead of time will disallow the update process to take place.

\section{Results}

Figure~\ref{fig:vulnerabilities} summarizes our results. All the collected
data are available in our GitHub repository at
\url{https://github.com/big-data-lab-team/container-vulnerabilities-paper}
with a Jupyter notebook to generate the figures.

\subsection{Detected Vulnerabilities}

\begin{figure*}
\includegraphics[width=\textwidth]{Figures/figure.pdf}
\caption{\label{fig:vulnerabilities} Number of vulnerabilities detected
by Anchore in container images. \textbf{(A)} Number of vulnerabilities
by container image and severity. \textbf{(B)} Combined
effect of image minification and package update on 5 container images. \textbf{(C, D)}
Number of vulnerabilities by number of packages, before and after
package update.}
\end{figure*}

A substantial number of vulnerabilities were found in the tested container
 images (Fig~\ref{fig:vulnerabilities}-\textbf{A}), with an average of 466 vulnerabilities
  per image and a median of 372. In comparison, no vulnerabilities were found in base
 Docker images \texttt{ubuntu:20.04} and \texttt{centos:7} after package
 update. Moreover, a significant fraction of detected vulnerabilities are
 of high severity
 (\href{https://www.first.org/cvss/specification-document}{CVSS} score
 >=7.0) and a few of them are of critical severity (CVSS >= 9.0). Remote
 attackers could possibly exploit these vulnerabilities to execute
 arbitrary code in the container, by crafting responses to specific network
 requests. The OS distribution is reported only for illustration purposes,
 as no obvious association between the OS distribution and numbers of
 vulnerabilities could be found.

Unsuprisingly, a strong linear relationship is found between the number of
detected vulnerabilities and the number of packages present in the
container image (Fig~\ref{fig:vulnerabilities}-\textbf{C}, r=0.83,
$p\textless10^{-10}$). On average, 1.8 vulnerabilities are introduced for
each new package installation. This observation motivates a systematic
review of software dependencies by application developers, to avoid
unnecessary packages in container images. This is also an argument in favor of lightweight
distributions such as Alpine. Compared to Ubuntu and Debian distributions,
CentOS images seem to have a lower number of vulnerabilities by package on
average, although data is too scarce to conclude.

\subsection{Effect of image udpate}
